---
title: "Reproducible Research - Course Project 1"
author: "Colette van Rooyen"
date: "02/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading and preprocessing the data

```{r}
        library(ggplot2)
        Data_Full <- read.csv("./Data/activity.csv")
        
        str(Data_Full)
```
The variables in this dataset are:  
1. **steps:** Number of steps taking in a 5-minute interval (missing values are coded as NA)  
2. **date:** The date on which the measurement was taken in YYYY-MM-DD format  
3. **interval:** Identifier for the 5-minute interval in which measurement was taken  

### What is mean total number of steps taken per day?  

1. Create a histogram  
2. Calculate the total number of steps taken per day    
3. Calculate and report the mean and median of the total number of steps taken per day  


```{r}
        steps_daily <- aggregate(Data_Full$steps, list(Data_Full$date), FUN = sum)
        colnames(steps_daily) <- c( "Date", "Steps")
        steps_daily
        
        # create histogram
        g <- ggplot(steps_daily, aes(Steps))
        g + geom_histogram(boundary=0, 
                           binwidth=2500, 
                           col="black", 
                           fill="yellow")+
                        ggtitle("Histogram of steps per day")+
                        xlab("Steps")+ylab("Frequency")+
                        theme(plot.title = element_text(face="bold",size=12))+
                        scale_x_continuous(breaks=seq(0,25000,2500))+
                        scale_y_continuous(breaks=seq(0,18,2))
        
        # Calculate and report the mean and median of the total number of steps taken per day
        mean(steps_daily$Steps, na.rm = TRUE)
        median(steps_daily$Steps, na.rm = TRUE)
```

