---
title: "Reproducible Research - Course Project 1"
author: "Colette van Rooyen"
date: "02/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading and preprocessing the data

```{r}
        library(ggplot2)
        Data_Full <- read.csv("./Data/activity.csv")
        
        str(Data_Full)
```
The variables in this dataset are:  
1. **steps:** Number of steps taking in a 5-minute interval (missing values are coded as NA)  
2. **date:** The date on which the measurement was taken in YYYY-MM-DD format  
3. **interval:** Identifier for the 5-minute interval in which measurement was taken  

### What is mean total number of steps taken per day?  

1. Create a histogram  
2. Calculate the total number of steps taken per day    
3. Calculate and report the mean and median of the total number of steps taken per day  


```{r}
        steps_daily <- aggregate(Data_Full$steps, list(Data_Full$date), FUN = sum)
        colnames(steps_daily) <- c( "Date", "Steps")
        steps_daily
        
        # create histogram
        g <- ggplot(steps_daily, aes(Steps))
        g + geom_histogram(boundary=0, 
                           binwidth=2500, 
                           col="black", 
                           fill="yellow")+
                        ggtitle("Histogram of steps per day")+
                        xlab("Steps")+ylab("Frequency")+
                        theme(plot.title = element_text(face="bold",size=12))+
                        scale_x_continuous(breaks=seq(0,25000,2500))+
                        scale_y_continuous(breaks=seq(0,18,2))
        
        # Calculate and report the mean and median of the total number of steps taken per day
        mean(steps_daily$Steps, na.rm = TRUE)
        median(steps_daily$Steps, na.rm = TRUE)
```
### What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  

```{r}
        average_daily_activity <- aggregate(Data_Full$steps, by=list(Data_Full$interval), FUN=mean, na.rm=TRUE)
        names(average_daily_activity) <- c("interval", "mean")
        plot(average_daily_activity$interval, average_daily_activity$mean, 
             type = "l", 
             col="red", 
             lwd = 2, 
             xlab="Interval", 
             ylab="Average number of steps", 
             main="Average number of steps per intervals")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
```{r}
        average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```
### Imputing missing values

1. Calculate and report the total number of missing values in the dataset  
2. Devise a strategy for filling in all of the missing values in the dataset.  
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.  
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
        sum(is.na(Data_Full$steps))
# Devise a strategy for filling in all of the missing values in the dataset.
        complete_steps <- average_daily_activity$mean[match(Data_Full$interval, average_daily_activity$interval)]
        str(complete_steps)
# Create a new dataset that is equal to the original dataset but with the missing data filled in.
        Data_Full_complete <- transform(Data_Full, steps = ifelse(is.na(Data_Full$steps), yes = complete_steps, no = Data_Full$steps))
        total_steps_complete <- aggregate(steps ~ date, Data_Full_complete, sum)
        names(total_steps_complete) <- c("date", "daily_steps")
        str(total_steps_complete)
# Make a histogram of the total number of steps taken each day
        hist(total_steps_complete$daily_steps, 
             col = "orange",
             xlab = "Total steps per day", 
             ylim = c(0,30), 
             main = "Total number of steps taken each day", 
             breaks = seq(0,25000,by=2500))
# Mean of the total number of steps taken per day
        mean(total_steps_complete$daily_steps)
# Median of the total number of steps taken per day
        median(total_steps_complete$daily_steps)
```


